### 2025-08-05 conteudo/methods/methods.tex
Review methods

The first method evaluates the proposed tool using an approach from the literature, 
comparing it against the BigCloneBench dataset \citep{bigclonebench}. 
The second method is an empirical analysis of a subset of functions within the 
AMD Display driver that our tool identified as duplicates. After completing these two 
methods, we triangulated the findings to analyze the results comprehensively. 
We chose triangulation because relying solely on the literature-based method has 
limitations and may be unreliable \citep{bigfail, litreview}. 
As a reminder, we have done all research approaches using the text similarity method 
as the code duplication detection method of the tool.

Given the validation of the tool’s capabilities, we proceed to understand the viability 
of mitigating the code duplication found by the tool. We conducted an ethnographic study 
using a participant-observation methodology to understand its viability. In this study, 
we aim to mitigate a subset of duplicated functions by refactoring them while interacting 
with the Linux community to gather feedback on our mitigation efforts.

The ethnographic study had two phases. In the first phase, the author explored the process 
of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
interact with the Linux community and get their feedback. In the second phase, we proposed 
to students in the course Free Software Development at the University of São Paulo to mitigate 
the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
pass the process of sending the mitigations as a PATCH.

\section{Evaluation of the Tool by the BigCloneBench dataset}

\label{sec:metbig}

The BigCloneBench dataset \citep{bigclonebench} is a Java code collection containing pairs of code duplications categorized by clone type, as defined in Section \ref{subsec:types}.
%
Following the methodology presented by \citep{tailor}, we sampled 20,000 pairs from each clone type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling approach to our tool to ensure a fair evaluation.

Unlike traditional methods, our tool does not distinguish between clone types and identifies duplications at the function level rather than the file level, which is typical in state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation purposes. Specifically, we considered every pair of functions with a similarity metric equal to or greater than a threshold \(X\) as duplicates and marked the corresponding file pairs. A correctly identified duplication pair counts as accurate for its clone type, while an incorrectly inferred pair is considered incorrect across all clone types. With these definitions, we computed the Recall metric from Section \ref{subsec:codemethods} for each clone type.
%
To understand the impact of varying the similarity threshold, we evaluated our tool using different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed and discussed the results.

@@

Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
an empirical study of the tool's inferred duplications.

@@

In this empirical analysis, we randomly sampled function pairs identified by the tool as duplicates. For each similarity threshold \(X\) (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%), we randomly selected ten function pairs with a similarity close to \(X\), allowing for a 1\% deviation. We implemented the random selection using the tool's source code.

Appendix \ref{app:emp} lists the randomly selected function pairs. We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool's effectiveness and understand the impact of the minimum similarity parameter.

\section{Ethnographic Study}

\label{sec:meteth}

Ethnography is a recognized qualitative empirical method for understanding people, 
their cultures, and their work practices \citep{bookethno}. It allows researchers 
to gain insights into values, beliefs, and ideas from the perspective of community 
members \citep{ethnosoft}. In our research, refactoring code duplications without 
community interaction would prevent us from understanding whether Linux maintainers 
would accept our approach or identify their concerns. Thus, we will use ethnography, 
positioning ourselves and students as Linux kernel contributors focused on removing 
identified  code duplications.

@@

For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

@@

A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to 
have students make a practical contribution to a free software project. In the first semester of 
2025, the professor teaching the course, who is also the advisor of this work, identified an 
opportunity to leverage the ArKanjo tool within this existing pedagogical structure. 
As a contribution option, the professor proposed that students mitigate code duplications found 
in the Linux kernel. The professor drove this initiative as a practical course activity, 
and the author of this thesis was not involved in using the tool to identify duplications or 
in developing approaches to mitigate them. The tasks were carried out entirely by the course 
staff and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of
two or three members. To assist students in this task, the professor and teaching assistants prepare multiple
alternatives for contributing to free software projects, including simpler contributions such as
updating documentation or fixing code style issues. In this context, two new options related to the
\texttt{anonymous-tool} were proposed alongside the others.

@@

For the second alternative, students were offered an experience similar to our initial participant
observation. In this model, the students were responsible for the entire workflow: executing the
ArKanjo on the AMD display driver, independently analyzing the results to identify a duplication
to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.

Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 
23 (16 undergraduate and 7 graduate)
formed 11 groups to pursue the first alternative. One graduate student also opted for the second 
alternative. Notably, none of these students had prior experience contributing to the Linux kernel, 
making this their first attempt to submit a patch as newcomers to the project.

>>>>

The first method evaluates the proposed tool using an approach from the literature, 
comparing it against the BigCloneBench dataset~\citep{bigclonebench}. 
The second method is an empirical analysis of a subset of functions within the 
AMD Display driver that our tool identified as duplicates. After completing these two 
methods, we triangulated the findings to analyze the results comprehensively. 
We chose triangulation because relying solely on the literature-based method has 
limitations and may be unreliable~\citep{bigfail,litreview}. 
As a reminder, we have done all research approaches using the text similarity method 
as the code duplication detection method of the tool.

Given the validation of the tool capabilities, we proceed to understand the viability 
of mitigating the code duplication found by the tool. We conducted an ethnographic study 
using a participant-observation methodology to understand its viability. In this study, 
we aim to mitigate a subset of duplicated functions by refactoring them while interacting 
with the Linux community to gather feedback on our mitigation efforts.

The ethnographic study had two phases. In the first phase, the author explored the process 
of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
interact with the Linux kernel community and get their feedback. In the second phase, we proposed 
to students in the course Free Software Development at the University of São Paulo to mitigate 
the code duplications found by the ArKanjo in the IIO subsystem and the AMD Display driver and 
pass the process of sending the mitigations as a PATCH.

\section{Evaluation of the Tool by the BigCloneBench dataset}
\label{sec:metbig}

The BigCloneBench dataset~\citep{bigclonebench} is a Java code collection containing pairs of code duplications categorized by clone type, as defined in Section~\ref{subsec:types}.
%
Following the methodology presented by~\citep{tailor}, we sampled 20,000 pairs from each clone type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling approach to our tool to ensure a fair evaluation.

Unlike traditional methods, our tool does not distinguish between clone types and identifies duplications at the function level rather than the file level, which is typical in state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation purposes. Specifically, we considered every pair of functions with a similarity metric equal to or greater than a threshold \(X\) as duplicates and marked the corresponding file pairs. A correctly identified duplication pair counts as accurate for its clone type, while an incorrectly inferred pair is considered incorrect across all clone types. With these definitions, we computed the Recall metric from Section~\ref{subsec:codemethods} for each clone type.
%
To understand the impact of varying the similarity threshold, we evaluated our tool using different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed and discussed the results.

@@

Due to the limitations of the BigCloneBench dataset, as noted by~\cite{bigfail}, and its lack of suitability for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with an empirical study of the tool inferred duplications.

@@

In this empirical analysis, we randomly sampled function pairs identified by the tool as duplicates. For each similarity threshold \(X\) (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%), we randomly selected ten function pairs with a similarity close to \(X\), allowing for a 1\% deviation. We implemented the random selection using the tool source code.

Appendix~\ref{app:emp} lists the randomly selected function pairs. We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool effectiveness and understand the impact of the minimum similarity parameter.

\section{Ethnographic Study}
\label{sec:meteth}

Ethnography is a recognized qualitative empirical method for understanding people, 
their cultures, and their work practices~\citep{bookethno}. It allows researchers 
to gain insights into values, beliefs, and ideas from the perspective of community 
members~\citep{ethnosoft}. In our research, refactoring code duplications without 
community interaction would prevent us from understanding whether Linux maintainers 
would accept our approach or identify their concerns. Thus, we will use ethnography, 
positioning ourselves and students as Linux kernel contributors focused on removing 
identified code duplications.

@@

For each duplicated function pair, for each function in the pair, we used the ArKanjo tool to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

@@

A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to 
have students make a practical contribution to a FLOSS project. In the first semester of 
2025, the professor teaching the course, who is also the advisor of this work, identified an 
opportunity to leverage the ArKanjo tool within this existing pedagogical structure. 
As a contribution option, the professor proposed that students mitigate code duplications found 
in the Linux kernel. The professor drove this initiative as a practical course activity, 
and the author of this work was not involved in using the tool to identify duplications or 
in developing approaches to mitigate them. The tasks were carried out entirely by the course 
staff and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of
two or three members. To assist students in this task, the professor and teaching assistants prepare multiple alternatives for contributing to FLOSS projects, including simpler contributions such as updating documentation or fixing code style issues. In this context, two new options related to the
Arkanjo tool were proposed alongside the others.

@@

For the second alternative, students were offered an experience similar to our initial participant
observation. In this model, the students were responsible for the entire workflow: executing the
ArKanjo on the AMD display driver, independently analyzing the results to identify a duplication
to fix, creating a patch to mitigate it, and sending the patch to the driver maintainers.

Of the students who chose to work on the ArKanjo-related tasks, 23 (16 undergraduate and 7 graduate)
formed 11 groups to pursue the first alternative. One graduate student also opted for the second 
alternative. Notably, none of these students had prior experience contributing to the Linux kernel, 
making this their first attempt to submit a patch as newcomers to the project.
